MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-06-16 latex/paper/sections/10-placement.tex
reformat results

In each \texttt{moveAll()}, we iterate through four mutually exclusive groups of placement objects: \texttt{DSPSiteInstCascades}, \texttt{CARRYSiteInstChains}, \texttt{RAMSiteInst}s, and \texttt{CLBSiteInst}s (single SLICEs that do not belong to CARRY chains).
For each placement object in the pass, we propose a move to a new placement on the device. 
If the proposed \texttt{Site} is already occupied by a resident or existing \texttt{SiteInst}, we evaluate a swap in their placements. 
Each move or swap is evaluated by calculating the HPWL cost of the placement before and after the proposed movement.
If the design has for example, 50 \texttt{DSPSiteInstCascades}, 100 \texttt{CARRYSiteInstChains}, 75 \texttt{RAMSiteInst}s, and 250 \texttt{CLBSiteInst}s, and we set a budget of 300 passses, then the algorithm will consider up to \(300 * (50 + 100 + 75 + 250) = 142500\) state transitions.

@@

If the move reduces the HPWL cost, then the movement is accepted outright.
If the move increases cost, then it can be accepted by chance if the global temperature is high enough to permit the hill-climb.
The temperature decreases with each pass until reaching zero, at which point, the algorithm makes exclusively energy-decreasing moves, effectively reducing into a greedy algorithm. 
The hope is that by this point, the placement will crystallize into a global optimum of minimum total HPWL.

\newcolumn
\begin{lstlisting}[language=java, caption={SA pseudocode: outer loop}, label={lst:sa_outer}]
public void placeDesign(PackedDesign packedDesign) {
    randomInitialPlacement(packedDesign);
    // init_temp=10000, alpha=0.98, max_passes=300
    initCoolingSchedule(10000, 0.98, 300)
    int passes = 0;
    while (passes < 300) {
        updateTemperature(passes);
        move(packedDesign)
        passes++;
    }
}

>>>>

In each \texttt{moveAll()}, we iterate through four mutually exclusive groups of placement objects: \texttt{DSPSiteInstCascades}, \texttt{CARRYSiteInstChains}, \texttt{RAMSiteInst}s, and \texttt{CLBSiteInst}s (single SLICEs that do not belong to CARRY chains).
For each placement object in the pass, we propose a new placement for that object on the device. 
Each proposed movement represents a proposed state transition of system. 
If the design has for example, 50 \texttt{DSPSiteInstCascades}, 100 \texttt{CARRYSiteInstChains}, 75 \texttt{RAMSiteInst}s, and 250 \texttt{CLBSiteInst}s, and we set a budget of 300 passses, then the algorithm will consider up to \(300 * (50 + 100 + 75 + 250) = 142500\) state transitions. 
If the proposed \texttt{Site} is already occupied by a resident or existing \texttt{SiteInst}, we evaluate a swap in their placements.

@@

If the move reduces the HPWL cost, then the movement is accepted outright.
If the move increases cost, then it can be accepted by chance if the global temperature is high enough to permit the hill-climb.
The temperature decreases with each pass until reaching zero, at which point, the algorithm makes exclusively energy-decreasing moves, effectively reducing into a greedy algorithm. 
The hope is that by this point, the placement has already crystallized into an approximate global optimum and the greediness of the algorithm will help the system settle deeper into the energy well faster. 
We can also switch from random selection to centroid selection when the temperature approaches near zero to speed up the convergence further.

\newcolumn
\begin{lstlisting}[language=java, caption={SA pseudocode: outer loop}, label={lst:sa_outer}]
public void placeDesign(PackedDesign packedDesign) {
    randomInitialPlacement(packedDesign);
    // init_temp=10000, alpha=0.98, max_passes=300
    initCoolingSchedule(10000, 0.98, 300)
    int passes = 0;
    while (passes < 300) {
        updateTemperature(passes);
        moveAll(packedDesign)
        passes++;
    }
}

### 2025-06-16 latex/paper/sections/11-placement_results.tex
placer comparison
reformat results

\end{multicols}

\subsection{Greedy Algorithm (Undirected/Random Moves)}
{
    \centering
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerGreedyRandom/random_placement.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerGreedyRandom/00000010.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerGreedyRandom/00000100.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerGreedyRandom/00000299.png}
    \captionof{figure}{Snapshots of placement of a 2048th order FIR Filter on a xc7z020 FPGA. \textbf{From left to right:} \textbf{(1)} Initial random placement, \textbf{(2)} After 10 passes, \textbf{(3)} After 100 passes, \textbf{(4)} Final placement after 300 passes.}
    \label{fig:PGRSnapshots}
}
{
    \centering
    \includegraphics[width=0.7\columnwidth]{figures/results/PlacerGreedyRandom/PlacerGreedyRandom_cost_history.png}
    \captionof{figure}{Total HPWL Cost vs number of passes. Final cost: 333013.}
    \label{fig:PGRCurve}
}

\subsection{Greedy Algorithm (Directed/Midpoint Moves)}
{
    \centering
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerGreedyMidpoint/random_placement.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerGreedyMidpoint/00000010.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerGreedyMidpoint/00000100.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerGreedyMidpoint/00000299.png}
    \captionof{figure}{Snapshots of placement of a 2048th order FIR Filter on a xc7z020 FPGA. \textbf{From left to right:} \textbf{(1)} Initial random placement, \textbf{(2)} After 10 passes, \textbf{(3)} After 100 passes, \textbf{(4)} Final placement after 300 passes.}
    \label{fig:PGMSnapshots}
}
{
    \centering
    \includegraphics[width=0.7\columnwidth]{figures/results/PlacerGreedyMidpoint/PlacerGreedyMidpoint_cost_history.png}
    \captionof{figure}{Total HPWL Cost vs number of passes. Final cost: 590578.}
    \label{fig:PGMCurve}
}

\subsection{Simulated Annealing (Undirected/Random Moves)}
{
    \centering
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealRandom/random_placement.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealRandom/00000010.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealRandom/00000100.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealRandom/00000299.png}
    \captionof{figure}{Snapshots of placement of a 2048th order FIR Filter on a xc7z020 FPGA. \textbf{From left to right:} \textbf{(1)} Initial random placement, \textbf{(2)} After 10 passes, \textbf{(3)} After 100 passes, \textbf{(4)} Final placement after 300 passes.}
    \label{fig:PARSnapshots}
}
{
    \centering
    \includegraphics[width=0.7\columnwidth]{figures/results/PlacerAnnealRandom/PlacerAnnealRandom_cost_history.png}
    \captionof{figure}{Total HPWL Cost vs number of passes. Final cost: 346676.}
    \label{fig:PARCurve}
}

\subsection{Simulated Annealing (Directed/Midpoint Moves)}
{
    \centering
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealMidpoint/random_placement.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealMidpoint/00000010.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealMidpoint/00000100.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealMidpoint/00000299.png}
    \captionof{figure}{Snapshots of placement of a 2048th order FIR Filter on a xc7z020 FPGA. \textbf{From left to right:} \textbf{(1)} Initial random placement, \textbf{(2)} After 10 passes, \textbf{(3)} After 100 passes, \textbf{(4)} Final placement after 300 passes.}
    \label{fig:PAMSnapshots}
}
{
    \centering
    \includegraphics[width=0.7\columnwidth]{figures/results/PlacerAnnealMidpoint/PlacerAnnealMidpoint_cost_history.png}
    \captionof{figure}{Total HPWL Cost vs number of passes. Final cost: 466855.}
    \label{fig:PAMCurve}
}

\subsection{Simulated Annealing (Hybrid: Random + Midpoint)}
{
    \centering
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealHybrid/random_placement.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealHybrid/00000010.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealHybrid/00000100.png}
    \includegraphics[valign=t, scale=0.25]{figures/results/PlacerAnnealHybrid/00000299.png}
    \captionof{figure}{Snapshots of placement of a 2048th order FIR Filter on a xc7z020 FPGA. \textbf{From left to right:} \textbf{(1)} Initial random placement, \textbf{(2)} After 10 passes, \textbf{(3)} After 100 passes, \textbf{(4)} Final placement after 300 passes.}
    \label{fig:PAHSnapshots}
}
{
    \centering
    \includegraphics[width=0.7\columnwidth]{figures/results/PlacerAnnealHybrid/PlacerAnnealHybrid_cost_history.png}
    \captionof{figure}{Total HPWL Cost vs number of passes. Final cost: 322933.}
    \label{fig:PAHCurve}
}

>>>>

Here, we present test results for a basic SA placer as well as test results for four simple variations as listed below. 
Note that midpoint simply means centroid.

\begin{itemize}
    \item \texttt{PlacerGreedyRandom}: all undirected random moves with greedy acceptance
    \item \texttt{PlacerGreedyMidpoint}: all directed centroid moves with greedy acceptance
    \item \texttt{PlacerAnnealRandom}: all undirected random moves with annealing acceptance \textbf{(the basic SA algorithm)}
    \item \texttt{PlacerAnnealMidpoint}: all directed centroid moves with annealing acceptance
    \item \texttt{PlacerAnnealHybrid}: 50\% random - 50\% centroid moves with annealing acceptance
\end{itemize}

{
    \centering
    \includegraphics[width=\columnwidth]{figures/results/combined_cost_history_linear.png}
    \captionof{figure}{All placers overlaid}
    \label{fig:placers_overlay}
}

\newcolumn
{
    \centering
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerGreedyRandom/random_placement.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerGreedyRandom/00000010.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerGreedyRandom/00000100.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerGreedyRandom/00000299.png}
    \captionof{figure}{PlacerGreedyRandom}
    \label{fig:PGRSnapshots}
}

\newcolumn
{
    \centering
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerGreedyMidpoint/random_placement.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerGreedyMidpoint/00000010.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerGreedyMidpoint/00000100.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerGreedyMidpoint/00000299.png}
    \captionof{figure}{PlacerGreedyMidpoint}
    \label{fig:PGMSnapshots}
}

{
    \centering
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealRandom/random_placement.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealRandom/00000010.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealRandom/00000100.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealRandom/00000299.png}
    \captionof{figure}{PlacerAnnealRandom}
    \label{fig:PARSnapshots}
}

\newcolumn
{
    \centering
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealMidpoint/random_placement.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealMidpoint/00000010.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealMidpoint/00000100.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealMidpoint/00000299.png}
    \captionof{figure}{PlacerAnnealMidpoint}
    \label{fig:PAMSnapshots}
}

{
    \centering
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealHybrid/random_placement.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealHybrid/00000010.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealHybrid/00000100.png}
    \includegraphics[valign=t, scale=0.13]{figures/results/PlacerAnnealHybrid/00000299.png}
    \captionof{figure}{PlacerAnnealHybrid}
    \label{fig:PAHSnapshots}
}
% {
%     \centering
%     \includegraphics[width=0.7\columnwidth]{figures/results/PlacerAnnealHybrid/PlacerAnnealHybrid_cost_history.png}
%     \captionof{figure}{Total HPWL Cost vs number of passes. Final cost: 322933.}
%     \label{fig:PAHCurve}
% }
